A Speculative Approach to Parallelization in Particle Swarm Optimization

### 2011-01-25 honors_thesis/si_paper/reviewer_response/response.tex

\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}

\begin{document}

\section*{Associate Editor Comments}

\subsection*{AE.1}

Reviewer comments: The quality of the article will be improved if you include a
review of existing parallelization methods for PSO, to produce a kind of
taxonomy. Then, also include comparisons with these approaches. You currently
compare only with a naive approach and adaptations of your approach.

Response: Copy and paste from Andrew's paper

NOT DONE

\subsection*{AE.2}

Reviewer comments: Please indicate if differences in performance are
statistically significant.

Response: Because all of the these finished before all of these, the t-test
gave \ldots

NOT DONE

\subsection*{AE.3}

Reviewer comments: It will also be good if you do make use of problems where
parallelization has benefit. For the functions that you use, there is not real
not for parallelization. Linked to this, when will parallelization be
beneficial?

Response: Get Andrew's paragraph that deals with this

NOT DONE

\subsection*{AE.4}

Reviewer comments: Please re-organize section 3: You start with a very long
theoretical discussion, and only then get to section 3.1.

Response: Good point.  We realized that we tried to cram too much into a single
section, so we split the two main parts of it into two separate sections.
There is now a section devoted entirely to the mathematics of the method and a
separate section describing its implementation.

NOT DONE

\section*{Reviewer 1 Comments}

\subsection*{1.1}

Reviewer comments: I'm a bit skeptical about the actual \emph{practical}
significance of the proposed method, [because it is difficult to know for any
given problem what the right swarm size is]. To be fair, this criticism applies
to most contributions focused on algorithms rather than on problems.

Response: This is a fair criticism, as it is true that experimentation with a
problem is required before one can know how best to solve it.  We added two
paragraphs to the end of section 2 to further make this point, making more
prominent reference to the No Free Lunch theorems.

\subsection*{1.2}

Reviewer comments: The organization and presentation of the ideas can be
greatly improved. The authors mix results and conclusions with the description
of the approach. This doesn't help the goal of the authors. A reader, cannot
appreciate the significance of some result or conclusion if he/she hasn't
understood what the authors are talking about. I recommend the authors to
restructure the whole paper but in particular Sections 3 and 4.

Response: This was not specific enough to use too much...

NOT DONE

\subsection*{1.3}

Reviewer comments: The notation used throughout the paper is confusing. In the
PSO literature, it is common to use x as a particles current position and some
other letters to denote a particle's personal best position and a particle's
neighborhood's best position. I know that the authors are using superscripts to
distinguish between these concepts but they don't really help.

Response: How do we fix this?

NOT DONE

\subsection*{1.4}

Reviewer comments: Using "rounds of evaluations" is confusing, and results
should be reported in terms of wall clock time.

Response: Rounds of evaluations correspond exactly to wall clock time, as we
mentioned.  However, we realize that the term is not a commonly used one, so we
changed "rounds of evaluations" to "time steps" in hopes of better conveying
our meaning. (Also incorporate Andrew's paragraph on why we're using time steps
instead of evaluations)

NOT DONE

\subsection*{1.5}

Reviewer comments: Since the authors are using Herrera et al.'s proposed
benchmarks and since they are using a large computing cluster, I don't see why
they can't use the same experimental protocol proposed by Herrera et al., which
is focused on large scale optimization. That protocol would better justify a
parallel algorithm because a 1000-dimensional problem is intuitively more
difficult than the 20-dimensional problems the authors are currently using. As
a side product, the authors can actually compare the results of the pruned
versions (which in my eyes, are new PSO variants) with those of other
state-of-the-art algorithms.

Response: We need to look up what he's talking about.

NOT DONE

\subsection*{1.6}

Reviewer comments: In the abstract, it is not clear the meaning of
"speculative". A reader without knowledge about parallel computing may not
understand what the authors mean.

Response: Need to address this.

NOT DONE

\subsection*{1.7}

Reviewer comments: The authors can cite Poli's paper about applications of PSO
algorithms in the first paragraph of the paper.

Response: We added the citation after the first sentence of the paper.

\subsection*{1.8}

Reviewer comments: What serves as the motivation for the authors's work is
insufficiently discussed.  The authors say that adding more particles leads to
"diminishing returns". This may be true when dealing with constant size
populations.  However, a recent trend is to vary the population size over time.
In fact, it has been shown that adding particles \emph{over time} may actually
solve the problem the authors refer to. Some references are "Incremental Social
Learning in Particle Swarms" by Montes de Oca and others published in IEEE SMCB
and "Efficient Population Utilization Strategy for Particle Swarm Optimizer" by
Hsieh and others in the same journal.

Response: See how papers are relevant (Andrew).  But also, assuming a fixed
computational budget, if at any point you have a smaller swarm than your budget
allows, you should do speculative evaluation.  That's our point.

NOT DONE

\subsection*{1.9}

Reviewer comments: In the first paragraph of Section 3, the authors say that
adding particles may not help past a certain swarm size. This is true is the
computation time allocated to the optimization process remains fixed. However,
in some problems, it is the solution quality what matters. In that case, adding
particles \emph{and} computational budget may indeed be the best thing to do.

Response: Most people have a fixed computational budget, function has a set
time.  That is the case we are interested in.

NOT DONE

\subsection*{1.10}

Reviewer comments: In Eq. 3. Why must the current position be worse than the
neighbor's position?

Response: Restate what the case means

NOT DONE

\subsection*{1.11}

Reviewer comments: One or two step-by-step examples would greatly improve
Section 3.

Response: We fixed section 3 by splitting it into two sections and reorganizing
some parts, making the examples that were there more clear.  Hopefully the
section is satisfactory now.

NOT DONE

\subsection*{1.12}

Reviewer comments: Isn't pruning against the idea of better using the available
computing resources? If one reduces the number of evaluated speculative
children, then fewer CPUs are going to be used.

Response: Pruning allows for the use of a larger swarm with the same number of
processors.  We tried to make this point more clear in the text.

NOT DONE

\subsection*{1.13}

Reviewer comments: Lines 29-30 of page 14. Stagnation is about the lack of
improvement of the best result. It has nothing to do with what the particles
are doing. I know that the PSO community confuses the two. I think this is a
nice opportunity to clarify the difference between search stagnation and the
state of the swarm.

Response: todo

NOT DONE

\subsection*{1.14}

Reviewer comments: All figures are not clear. The authors should find line
styles that are more distinguishable.

Response: ...

NOT DONE

\subsection*{1.15}

Reviewer comments: Page 27, line 40. Algorithms are suited for problems not the
other way around.

Response: ...

NOT DONE

\subsection*{1.16}

Reviewer comments: Page 31. Line 46. Molia -> Molina. Is it a technical report?
Which number?  Isn't it published with the special issue?

Response: ...

NOT DONE

\section*{Reviewer 2 Comments}

I found the idea very interesting and the paper well-written. However, there
are still some points that need further investigation.  As far as I am
concerned, the paper has merit and shall be accepted after a revision that will
address these remarks.

\subsection*{2.1}

Reviewer comments: How does it scale with the problem's dimension? There is
experimental evidence that the required swarm size for achieving a prescribed
accuracy in a specific problem, increases almost exponentially with its
dimension. The proposed idea can make this necessity milder. Is there any
evidence on how well it scales when dimension increases? (e.g. probably the
increase in the required number of particles is closer to the linear rather
than the exponential case).

Response: ...

NOT DONE

\subsection*{2.2}

Reviewer comments: In the first sections of the paper it is argued that the
proposed "...method can easily be extended to arbitrary topologies". However,
in later sections it is apparent that this may be possible if the underlying
neighborhood topology is not "dense" enough to cause heavy communication
traffic to the processors.  Can you be more precise to this and quantify it?
For example, given a swarm size and a common network communication is there any
identifiable limit to the neighborhood topology such that the algorithm retains
an acceptable performance?

Response: ...

NOT DONE

\subsection*{2.3}

Reviewer comments: My personal experience with the PSO versions used in the
paper suggests that, in large dimensions, it is preferrable to use larger
swarms for less iterations rather than small swarms for more iterations. This
is in contrast to the argued behavior of PSO for the Sphere function (Section
1).  Did you observe the same behavior in other functions?  Could you provide
some evidence on this because, due to the swarm's diversity loss, it seems
somehow counterintuitive.

Response: ...

NOT DONE

\section*{Reviewer 3 Comments}

\subsection*{3.1}

Reviewer comments: There are too many redundancy sentences declaring this paper
is not proposing a new algorithm in the introduction.

Response: ...

NOT DONE

\subsection*{3.2}

Reviewer comments: The velocity updating formular is different from the
original PSO algorithm.  The weighting factor, X, should only affects the old
velocity, not all the items.

Response: reference constricted PSO in formulas

NOT DONE

\subsection*{3.3}

Reviewer comments: Some reference are not completed. For example, the
journal/conference name of the 4th reference is missing. Please check it
carefully.

Response: ...

NOT DONE

\subsection*{3.4}

Reviewer comments: Please increasing the reference to extend the integrity of
the background.

Response: ...

  a)Pei-Wei Tsai, Jeng-Shyang Pan, Shyi-Ming Chen, Bin-Yih Liao, and Szu-Ping
Hao, 2008, "Parallel Cat Swarm Optimization," Proceedings of the Seventh
International Conference on Machine Learning and Cybernetics, Kunming, China,
Jul. 12-15, 2008, pp. 3328-3333.

  b)Jui-Fang Chang, Shu-Chuan Chu, John F. Roddick and Jeng-Shyang Pan, 2005,
"A Parallel Particle Swarm Optimization Algorithm with Communication
Strategies", Journal of Information Science and Engineering, Vol. 21, No. 4,
pp. 809-818.

NOT DONE

\end{document}

### 2011-01-25 honors_thesis/si_paper/paper.tex
A few revisions to the si paper to address reviewer comments

Particle swarm optimization (PSO) has been found to be a highly robust and
effective algorithm for solving many types of optimization problems.  For much
of the algorithm's history, PSO was run serially on a single machine.  However,
the world's computing power is increasingly coming from large clusters of
processors.  In order to efficiently utilize these resources for
computationally intensive problems, PSO needs to run in parallel.

@@

The balance of this paper is organized as follows. \sec{pso} describes the
particle swarm optimization algorithm.  \sec{sepso} describes how speculative
evaluation can be done in parallel PSO to perform two iterations at once.  In
\sec{relax}, we discuss various methods of improving the performance of
speculative evaluation in PSO, all of which break the requirement of strictly
reproducing the behavior of the original algorithm.  \sec{setup} describes the
experiments we ran, and \sec{results} presents the results of those
experiments.  In \sec{conclusion} and \sec{future} we conclude and discuss
future work.

@@

@@

It is not trivial in some parallel architectures to determine which speculative
position was the correct next position of each particle.  In the rest of this
section we move from the mathematics of our method to its implementation.
First we discuss the relatively easy case of a centralized parallel PSO
algorithm with a master computer and many slaves.  In such an architecture, the
master keeps track of all necessary information with only trivial message
passing needed.  Then we discuss the more complicated case of a distributed
algorithm, where each particle is on its own and needs to send and receive
messages to and from other particles.  Finally we discuss the further
complications of a dynamic topology, where a particle's neighbors change from
one iteration to another.

>>>>

Particle swarm optimization (PSO) has been found to be a highly robust and
effective algorithm for solving many types of optimization
problems~\cite{poli-2008-pso-applications}.  For much of the algorithm's
history, PSO was run serially on a single machine.  However, the world's
computing power is increasingly coming from large clusters of processors.  In
order to efficiently utilize these resources for computationally intensive
problems, PSO needs to run in parallel.

@@

The balance of this paper is organized as follows. \sec{pso} describes the
particle swarm optimization algorithm.  \sec{sepso} describes mathematically
how speculative evaluation can be done in parallel PSO to perform two
iterations at once, and \sec{implementation} gives information on how to
implement this method.  In \sec{relax}, we discuss various methods of improving
the performance of speculative evaluation in PSO, all of which break the
requirement of strictly reproducing the behavior of the original algorithm.
\sec{setup} describes the experiments we ran, and \sec{results} presents the
results of those experiments.  In \sec{conclusion} and \sec{future} we conclude
and discuss future work.

@@

The ideal topology and swarm size for PSO depend on the objective function.
Researchers have devised various benchmark functions and have found that the
ideal topology for one function may perform very poorly for another function.
The No Free Lunch Theorems for Optimization show that this is true in
general---if an algorithm performs well on average for one class of functions
then it must do poorly on average for other
problems~\cite{wolpert-1997-nfl-for-optimization}.

An attempt to standardize PSO found that although a Complete swarm of 50
particles converged to the global optimum more quickly for many benchmark
functions than the same swarm size with a Ring topology, it was also more
likely to permaturely converge to local optima for other functions.  The study
found no significant improvement for any other swarm size between 20 and 100
and concluded with a recommendation to use a swarm of 50 particles with a Ring
topology as a starting point.  The authors acknowledged that choosing the ideal
topology requires thorough experimentation for the particular
problem~\cite{bratton-2007-defining-a-standard-for-pso}.  Other authors have
explained that a large swarm with a sparse topology propogates information
slowly~\cite{montes-de-oca-2009-frankensteins-pso}.

@@

\section{Implementing Speculative Evaluation}
\label{sec:implementation}

It is not trivial in some parallel architectures to determine which speculative
position was the correct next position of each particle.  In this section we
move from the mathematics of our method to its implementation.  First we
discuss the relatively easy case of a centralized parallel PSO algorithm with a
master computer and many slaves.  In such an architecture, the master keeps
track of all necessary information with only trivial message passing needed.
Then we discuss the more complicated case of a distributed algorithm, where
each particle is on its own and needs to send and receive messages to and from
other particles.  Finally we discuss the further complications of a dynamic
topology, where a particle's neighbors change from one iteration to another.
